Texture mapping 3D planar models of indoor environments with noisy camera poses

### 2013-01-18 spie/paper.tex
new paper

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\begin{abstract}

  Automated 3D modeling of building interiors is used in applications
  such as virtual reality and environment mapping. Texturing these
  models allows for photorealistic visualizations of the data
  collected by such modeling systems. Camera poses obtained by these
  systems and used for texturing often suffer from inaccuracies
  however, resulting in visible discontinuities when successive images
  are projected adjacently onto a surface for texturing. Existing
  methods to stitch images together are often computationally
  expensive and work independently of pose estimates and geometry
  information. We propose a method to refine camera poses using both
  existing estimates and geometry information, followed by two
  different methods to composite images together, based on the
  uniformity of available images. The effectiveness of our methods is
  demonstrated on a number of different indoor environments.
\end{abstract}

@@

The first step in automated 3D modeling is the physical scanning of an
environment's geometry. An indoor modeling system must recover its own
poses within an environment while simultaneously reconstructing the 3D
structure of the environment itself. \cite{chen2010indoor,
  liu2010indoor, kua2012loopclosure}. This is known as the
simultaneous localization and mapping (SLAM) problem, and is generally
solved by taking readings from laser range scanners, cameras, and
inertial measurement units (IMUs) at multiple locations within the
environment.

In this paper, we work with data obtained from a backpack-mounted
system, carried by an ambulatory human. Such a system provides
advantages over more common wheel-mounted systems in terms of agility
and portability, but results in much higher localization error
\cite{liu2010indoor}. As a result, common methods for texture mapping
generally produce poor results, leading to the development of the
approaches contained in this paper. Besides localization information
and images, environment geometry is required as well for texture
mapping. In this paper we work with models obtained by fitting planar
surfaces to point clouds generated by the backpack system
\cite{sanchez2012point}.A.

Though our texture mapping procedure was designed with the
aforementioned system in mind, it is applicable to other systems,
provided the required inputs of planar geometry, images with rough
extrinsic pose estimates, and either known or estimated intrinsic
camera calibration matrices. An important benefit of our method is its
relatively low complexity and runtime, and independence of many steps,
allowing for live tuning of parameters and easy accomodation of new
input data given updates in localization or model-generation
algorithms.

@@

@@

\begin{figure}
  \begin{minipage}[b]{0.45\linewidth}
    \centering
    \includegraphics[height=2in]{Projection.pdf}
    \caption{Planes are specified in 3D space by four corners $C_1$ to
      $C_4$. Images are related to each plane through the camera
      matrics $P_{1..m}$. }
    \label{fig:projection}
  \end{minipage}
  \hspace{0.5cm}
  \begin{minipage}[b]{0.45\linewidth}
    \centering
    \includegraphics[height=1.5in]{scoringFunction.jpg}
    \caption{We minimize camera angle $\alpha$ and distance $d$ by
      maximizing the scoring function $\frac{1}{d} (-1 \cdot \vec{c})
      \cdot \vec{n}$}
    \label{fig:scoringFunction}
  \end{minipage}
\end{figure}

@@

We choose to work with rectangular units to ensure that borders
between any two distinct images in the final texture are either
horizontal or vertical. Since most strong environmental features
inside buildings are horizontal or vertical, any visible seams in our
texture intersect them minimally and are less noticeable.

@@

\begin{figure}
  \centering \subfloat[][]{\includegraphics[width=3in,
    height=1.2in]{wall2_naive.jpg}}
  \subfloat[][]{\includegraphics[width=3in,
    height=1.2in]{wall2_naive_shift.jpg}}

  \centering \subfloat[][]{\includegraphics[width=3in,
    height=1.2in]{wall2_cache_shift.jpg}}
  \subfloat[][]{\includegraphics[width=3in,
    height=1.2in]{wall2_shortest_shift.pdf}}

  \centering \subfloat[][]{\includegraphics[width=3in,
    height=1.2in]{wall2_cache_shift_blend.pdf}}
  \subfloat[][]{\includegraphics[width=3in,
    height=1.2in]{wall2_shortest_shift_blend.pdf}}
  \caption{(a) Tile-based texturing. (b) Tile-based texturing after
    image alignment. (c) Tile-based texturing after image alignment
    with caching. (d) Shortest path texturing after image
    alignment). (e) Same as (c) with blending. (f) Same as (d) with
    blending.}
  \label{fig:compareAll}
\end{figure}

@@

Image projections can also be aligned by iteratively adjusting camera
poses to maximize matches. This is generally done over 6 degrees of
freedom \cite{liu2010indoor}. The approach in the cited paper works
well locally, but over larger areas results again in error
accumulation and drift, as shown in Figure
\ref{fig:mosaic3D}(b). Furthermore, it adjusts image locations using
a long iterative process, which results in high runtime.

@@

\begin{figure}
  \centering
  \subfloat[][]{\includegraphics[height=1.5in]{geometryAlign_planes.png}}
  \subfloat[][]{\includegraphics[height=1.5in]{geometryAlignment.jpg}}
  \caption{(a) When texturing the red surface, it makes sense to align
    images to surface boundaries and intersections with other
    surfaces. (b) Red lines are geometry-based lines, while green
    lines are detected in the image via Hough transform. Above is the
    original projection using the input camera poses, and below is the
    projection after being adjusted for alignment.}
  \label{fig:geometryAlignment}
\end{figure}

@@

The result of this caching approach is shown in Figure
\ref{fig:compareAll}(c), where seams are now reduced as
compared to Figure \ref{fig:compareAll}(b).

@@

With this completed, we have now mapped every location on the plane to
at least one image, and have minimized the number of images, as well
as the discontinuities at their borders. As seen in Figure
\ref{fig:compareAll}(d), this shortest path method has fewer visible
discontinuities than Figure \ref{fig:compareAll}(c) corresponding to
the tile caching approach\footnote{In Figure \ref{fig:compareAll}(d),
  we arbitrarily chose one image for texturing where images overlap,
  as blending will be discussed in section \ref{sec:blending}.}. This
is especially evident when comparing the posters in the images, which
have clear misalignment over seams in Figure \ref{fig:compareAll}(c),
but are much more aligned in Figure \ref{fig:compareAll}(d). This
shortest path approach approach directly reduces the cost of each
image boundary, while the tile caching method uses a scoring function
that only approximates this effect. Furthermore, this approach
guarantees the best selection of images to minimize seams, while the
sequential tile caching method may select images early on that turn
out to be poor choices once subsequent tiles have been processed. This
shortest path approach is also far less intensive in terms of memory
usage and runtime, both during texture generation and rendering, as it
does not require discretizing planes or images.

@@

After performing linear alpha blending across overlapping regions, our
texture mapping process is complete. Figures \ref{fig:compareAll}(e)
and \ref{fig:compareAll}(f) show the blended versions of Figures
\ref{fig:compareAll}(c) and \ref{fig:compareAll}(d) respectively. It
is clear that the shortest path approach exhibits much better
alignment and fewer seams than the tile-caching method, as Figure
\ref{fig:compareAll}(f) clearly has the best visual quality among the
textures in Figure \ref{fig:compareAll}.

\section{Results and Conclusions}
\label{sec:resultsAndConclusions}
Examples of ceilings and floors textured with the tile caching
approach, and walls textured with the shortest path approach, are
displayed in Figure \ref{fig:results}. High resolution texture
comparisons, as well as further examples and interactive walkthroughs are
available at \footnote{\url{http://www.eecs.berkeley.edu/~pcheng/textureMapping}}.

As mentioned earlier, our approach is quite efficient. The top wall in
Figure \ref{fig:results}(a) was generated with 7543 $\times$ 776
pixels, and spans a 40-meter long wall. Given 41000 input images, a
2.8GHz consumer-grade laptop takes approximately a minute to pick 36
candidate images, followed by another minute to perform both image
alignment and the shortest path texturing method, with the majority of
computation time spent calculating SIFT matches within the SiftGPU
framework. While not quite real-time visualization, the process is
capable of generating quick updates after changes in various
parameters or modifications to input data.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\begin{abstract}
  Automated 3D modeling of building interiors is used in applications
  such as virtual reality and environment mapping. Texturing these
  models allows for photorealistic visualizations of the data
  collected by such modeling systems. Camera poses obtained by these
  systems and used for texturing often suffer from inaccuracies
  however, resulting in visible discontinuities when successive images
  are projected onto a surface for texturing. Existing methods to
  stitch images together are often computationally expensive and work
  independently of pose estimates and geometry information. We propose
  an efficient method to refine camera poses in 2D using both existing
  estimates and geometry information, followed by two different
  methods to composite images together, based on the geometry of
  surfaces being textured. The effectiveness of our methods is
  demonstrated on a number of different indoor environments.
\end{abstract}

@@

The first step in automated 3D modeling is the physical scanning of an
environment's geometry. An indoor modeling system must recover its own
poses within an environment while simultaneously reconstructing the 3D
structure of the environment itself \cite{chen2010indoor,
  liu2010indoor, kua2012loopclosure}. This is known as the
simultaneous localization and mapping (SLAM) problem, and is generally
solved by taking readings from laser range scanners, cameras, and
inertial measurement units (IMUs) at multiple locations within the
environment.

In this paper, we work with data obtained from a backpack-mounted
system, carried by an ambulatory human. Such a system provides
advantages over more common wheel-mounted systems in terms of agility
and portability, but results in higher difficulty and error when
performing localization \cite{liu2010indoor}. As a result, approaches
to texture mapping that rely on accurate camera poses produce poor
results, leading to the development of the methods contained in this
paper. Besides localization information and images, environment
geometry is required as well for texture mapping. In this paper we
work with low-resolution models obtained by fitting planar surfaces to
point clouds generated by the backpack system \cite{sanchez2012point}.

SOME CONCLUSION TO THE INTRODUCTION GOES HERE Though our texture
mapping procedure was designed with the aforementioned system in mind,
it is applicable to other systems, provided the required inputs of
planar geometry, images with rough extrinsic pose estimates, and
either known or estimated intrinsic camera calibration matrices. An
important benefit of our method is its relatively low complexity and
runtime, and independence of many steps, allowing for efficient tuning
of parameters and easy accomodation of new input data given updates in
localization or model-generation algorithms.

@@

\section{Background}
\label{sec:background}
\subsection{Existing Approaches}
\label{sec:existingApproaches}
\subsection{

@@

  \begin{figure}
    \begin{minipage}[b]{0.45\linewidth}
      \centering
      \includegraphics[height=1.5in]{Projection.pdf}
      \caption{Planes are specified in 3D space by four corners $C_1$
        to $C_4$. Images are related to each plane through the camera
        matrics $P_{1..m}$. }
      \label{fig:projection}
    \end{minipage}
    \hspace{0.5cm}
    \begin{minipage}[b]{0.45\linewidth}
      \centering
      \includegraphics[height=1.125in]{scoringFunction.jpg}
      \caption{We minimize camera angle $\alpha$ and distance $d$ by
        maximizing the scoring function $\frac{1}{d} (-1 \cdot
        \vec{c}) \cdot \vec{n}$}
      \label{fig:scoringFunction}
    \end{minipage}
  \end{figure}

@@

  We choose to work with rectangular units to ensure that borders
  between any two distinct images in the final texture are either
  horizontal or vertical. Since most strong environmental features
  inside buildings are horizontal or vertical, any visible seams in
  our texture thus intersect them minimally and are less noticeable.

@@

\begin{figure}
  \centering \subfloat[][]{\includegraphics[width=3.4in,
    height=1.2in]{wall2_naive.jpg}}
  \subfloat[][]{\includegraphics[width=3.4in,
    height=1.2in]{wall2_naive_shift.jpg}}

  \centering \subfloat[][]{\includegraphics[width=3.4in,
    height=1.2in]{wall2_cache_shift.jpg}}
  \subfloat[][]{\includegraphics[width=3.4in,
    height=1.2in]{wall2_shortest_shift.pdf}}

  \centering \subfloat[][]{\includegraphics[width=3.4in,
    height=1.2in]{wall2_cache_shift_details.png}}
  \subfloat[][]{\includegraphics[width=3.4in,
    height=1.2in]{wall2_shortest_shift_details.png}}

  \centering \subfloat[][]{\includegraphics[width=3.4in,
    height=1.2in]{wall2_cache_shift_blend.pdf}}
  \subfloat[][]{\includegraphics[width=3.4in,
    height=1.2in]{wall2_shortest_shift_blend.pdf}}

  \caption{(a): Tile-based texturing. (b): Tile-based texturing after
    image alignment. (c): Tile-based texturing after image alignment
    with caching. (d): Shortest path texturing after image
    alignment). (e,f): Comparison of image artifacts in (c)
    vs. (d). (g,h): Blending applied to (c) and (d).}
  \label{fig:compareAll}
\end{figure}

@@

Image projections can also be aligned by iteratively adjusting camera
poses to maximize matches. This is generally done over 6 degrees of
freedom \cite{liu2010indoor}. The approach in [2] works well locally,
but over larger areas results again in error accumulation and drift,
as shown in Figure \ref{fig:mosaic3D}(b). Furthermore, it adjusts
image locations using a long iterative process, which results in high
runtime.

@@

\begin{figure}
  \centering
  \subfloat[][]{\includegraphics[height=1.5in]{geometryAlign_planes.png}}
  \subfloat[][]{\includegraphics[height=1.5in]{geomAlign}}
  \caption{(a) When texturing the red surface, it makes sense to align
    images to surface boundaries and intersections with other
    surfaces. (b) The darker lines are geometry-based lines, while
    lighter lines are lines detected in the image via Hough
    transform. Above is the original projection using the input camera
    poses, and below is the projection after rotation and translation
    for alignment.}
  \label{fig:geometryAlignment}
\end{figure}

@@

The result of this caching approach is shown in Figure
\ref{fig:compareAll}(c), where seams are now reduced as compared to
Figure \ref{fig:compareAll}(b). However, some discontinuities are
still present, as visible in the many posters on the wall with breaks
in their borders.

@@

With this completed, we have now mapped every location on the plane to
at least one image, and have minimized the number of images, as well
as the discontinuities at their borders. As seen in Figure
\ref{fig:compareAll}(d), this shortest path method has fewer visible
discontinuities than Figure \ref{fig:compareAll}(c) corresponding to
the tile caching approach\footnote{In Figure \ref{fig:compareAll}(d),
  we arbitrarily chose one image for texturing where images overlap,
  as blending will be discussed in section \ref{sec:blending}.}. This
is especially evident when comparing the posters in the images, as
seen in Figures \ref{fig:compareAll}(e) and
\ref{fig:compareAll}(f). This shortest path approach approach directly
reduces the cost of each image boundary, while the tile caching method
uses a scoring function that only approximates this
effect. Furthermore, this approach guarantees the best selection of
images to minimize seams, while the sequential tile caching method may
select images early on that turn out to be poor choices once
subsequent tiles have been processed. This shortest path approach is
also far less intensive in terms of memory usage and runtime, both
during texture generation and rendering, as it does not require
discretizing planes or images.

@@

After performing linear alpha blending across overlapping regions, our
texture mapping process is complete. Figures \ref{fig:compareAll}(g)
and \ref{fig:compareAll}(h) show the blended versions of Figures
\ref{fig:compareAll}(c) and \ref{fig:compareAll}(d) respectively. It
is clear that the shortest path approach exhibits much better
alignment and fewer seams than the tile-caching method, as Figure
\ref{fig:compareAll}(h) clearly has the best visual quality among the
textures in Figure \ref{fig:compareAll}.

\section{Results and Conclusions}
\label{sec:resultsAndConclusions}
Examples of ceilings and floors textured with the tile caching
approach, and walls textured with the shortest path approach, are
displayed in Figure \ref{fig:results}. High resolution colored texture
comparisons, as well as further examples and interactive walkthroughs
are available at
\footnote{\url{http://www.eecs.berkeley.edu/~pcheng/indoormapping}}.

As mentioned earlier, our approach is quite efficient. The top wall in
Figure \ref{fig:results}(a) was generated with 7543 $\times$ 776
pixels, and spans a 40-meter long wall. Given 41000 input images, a
2.8GHz consumer-grade laptop takes approximately a minute to pick 36
candidate images, followed by another minute to perform both image
alignment and the shortest path texturing method, though over 75\% of
that time is spent calculating SIFT matches within the SiftGPU
framework. While not real-time, the process is capable of generating
quick updates after changes in various parameters or modifications to
input data, and if integrated directly into a modeling system, could
provide live visualization as data is collected.
